Not `Won't you be my neighbor?' \\ But `Should you be my neighbor?'

### 2015-01-23 EBrush_notes_flocking_evolution.tex
im about to go through correlations and H2 more carefully

\section{Model }
\subsection{Learning dynamics }
An individual's opinion changes with a probability that depends on its social interactions and the strength of the environmental stimulus. Specifically, an individual's opinion $\sigma_i$ can be either $1$ or $-1$ and $\sigma_i$ switches with probability $w_i(\sigma_i)=\frac{1}{2}(-\sigma_i\sum_jA_{ij}\sigma_j )(1-\beta_i\sigma_i)$ where $A_{ij}$ indicates the strength of the connection from $j$ to $i$ and $\beta_i$ indicates the strength of the environmental stimulus perceived by $i$.  The stochastic dynamics of the opinions can be written down with a master equation \cite{Glauber:1963fk}.  The vector $\vec{v}(t)$ of expected values, $q_i(t)=\langle \sigma_i(t)\rangle$, satisfies the following differential equations: \cite{Glauber:1963fk}
\begin{equation*}
\frac{d\vec{v}}{dt}=M\vec{v}+\vec{\beta}^T(\vec{1}-\vec{v}).
\end{equation*}
We simplify this by writing $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}^T\vec{1}$, where $A=M-B$ and $B$ is a diagonal matrix with the values of $\vec{\beta}$ along the diagonal. In the absence of the environmental signal, there is no trend in the dynamics. In the presence of the environmental signal, however, those individuals who perceive the signal will be more likely to switch to having an opinion of $1$ so their expected value will increase, and consequently so will the values of the those paying attention to the informed individuals. Thus, having a higher expected value can be interpreted as having more information about the information and having a lower expected value can be interpreted as having less information.

@@

\subsection{ESS strategies }
Surprisingly, even without imposing costs on paying attention to more neighbors, paying attention to the whole flock is not usually the best strategy. The ESS strategy of the number of neighbors to pay attention strongly depends on which selection pressure is being applied.  If a focal bird pays attention to many others, its opinion will be similar to theirs. This will prevent it from being the most informed and hence the first to locate resources. Thus, the uninvadable strategy is to pay attention to few neighbors when selection is due to knowledge about resources. On the other hand, having many neighbors will prevent the focal bird from being the least informed and thus the most likely to be eaten.  Thus, the uninvadable strategy is to pay attention to many neighbors when selection is due to knowledge about predators. Figure \ref{ESS} shows the ESS strategy when selection is due to knowledge about resources is always lower than the ESS strategy when selection is due to knowledge about predators (Figure \ref{ESS}).

The ESS strategies when selection is due to knowledge about predators show two interesting features. First, there are many parameters for which there are multiple ESS strategies. It might seem that if a strategy outperforms all others when it is the resident strategy, it should also outperform other strategies as an invader. However, when there are multiple ESS strategies, if we take any two of those ESS strategies, each is uninvadable by the other. This indicates how important the social environment set by the resident strategy is. The resident strategy essentially sets a level of background noise. As an example, consider the strategies of having sixteen or seventeen neighbors when the radius of the signal is $r=0.1$. Figure \ref{ESS} shows how the probability of being predated for each of these invasion scenarios depends on the number of birds in the group and the number of informed neighbors the invader has.  A bird with seventeen neighbors will have more uninformed neighbors than one with sixteen neighbors and thus will have less information itself. This poor information is  costly when the resident strategy is to have sixteen neighbors since the residents will have less noisy information, which prevents an mutant with seventeen neighbors to invade. Figure \ref{ESS} shows that by paying attention to more uninformed neighbors this mutant is likely to be eaten when there are many truly informed birds. However, having less noisy information does not allow a sixteen neighbor mutant to invade a seventeen neighbor group since, regardless of the number of uninformed neighbors the invader has, the background level of noise is much higher in a group all of whom have more uninformed neighbors. The sixteen neighbor mutant also pays a cost by being less likely to pay attention to the right neighbors when there are very few truly informed birds.  Figure \ref{ESS} shows that by being less likely to pay attention to the right neighbors this mutant is likely to be eaten when there are few truly informed birds.

These opposing forces caused by the costs of paying attention to more uninformed neighbors and the costs of not paying attention to informed neighbors also explains the second feature of the ESS strategies: they are non-monotonic with respect to the the radius of the signal, i.e. how public the signal is (Figure \ref{ESS}). Figure \ref{ESS} shows that there are essentially three categories of radii: When the radius is low, the signal is private, with only a few birds having true information. When the radius is intermediate, the signal is public, with many birds having true information. When the radius is high, the signal is universal, with essentially all birds in the flock having true information. When the signal is private, paying attention to too few birds is costly because with more neighbors it is more likely to have a truly informed neighbor, but paying attention to too many uninformed birds can also be costly. This leads to a high but not maximal ESS strategy. When the signal is public, the costs of paying attention to uninformed birds are reduced and the ESS strategy increases to become as high as possible. When the signal is universal, not paying attention to informed birds does not impose a cost. Additionally, in a group in which the residents pay attention to all of their neighbors, any mutant strategy can get just as much information by paying attention to fewer neighbors so the maximal strategy is no longer an ESS (Figure \ref{ESS}).

@@

\section{Discussion}
explore vs. exploit 
risk aversion vs. risk tolerance

expect single ESS strategies but social information depends on the social environment set by the resident strategy
expect ESS strategies to decrease as the signal becomes more public, but we find the opposite

\section{Figures }
\begin{figure}[ht]
\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/ESSfigure.pdf}
\caption{\label{ESS} The ESS number of neighbors is always higher if selection is due to predation than due to resources. When selection is due to predation, there can be multiple ESS strategies, which are a non-monotonic function of the radius of the signal. In (a), we show the (possibly multiple) ESS strategies as a function of the radius of the signal. In (b) we show the probability of there being a given number of informed birds for signals with different radii. In (c) and (d), each heat map shows the probability that an invader is eaten as a function of the number of informed birds in the flock on the horizontal axis and the number of informed neighbors the invader has on the vertical axis. In (c), the resident strategy is to have $16$ neighbors and the invader strategy is to have $17$ neighbors. In (d), the resident strategy is to have $17$ neighbors and the invader strategy is to have $16$ neighbors. Parameters:  in all panels $\beta=1$, $T=1$, in (c) and (d) $r=0.1$. 
}
\end{figure}

>>>>

\section{Model }
\subsection{Learning dynamics }
An individual's opinion changes with a probability that depends on its social interactions and the strength of the environmental stimulus. Specifically, an individual's opinion $\sigma_i$ can be either $1$ or $-1$ and $\sigma_i$ switches with probability $w_i(\sigma_i)=\frac{1}{2}(-\sigma_i\sum_jA_{ij}\sigma_j )(1-\beta_i\sigma_i)$ where $A_{ij}$ indicates the strength of the connection from $j$ to $i$ and $\beta_i$ indicates the strength of the environmental stimulus perceived by $i$.  The stochastic dynamics of the opinions can be written down with a master equation \cite{Glauber:1963fk}.  The vector $\vec{v}(t)$ of expected values, $q_i(t)=\langle \sigma_i(t)\rangle$, satisfies the following differential equations: \cite{Glauber:1963fk}
\begin{equation*}
\frac{d\vec{v}}{dt}=M\vec{v}+\vec{\beta}^T(\vec{1}-\vec{v}).
\end{equation*}
We simplify this by writing $\frac{d\vec{v}}{dt}=A\vec{v}+\vec{\beta}^T\vec{1}$, where $A=M-B$ and $B$ is a diagonal matrix with the values of $\vec{\beta}$ along the diagonal.

These stochastic dynamics will lead to an equilibrium probability distribution over sets of opinions that agrees with the Boltzmann equilibrium distribution for a system with energy or Hamiltonian given by 
\begin{equation*}
E(\vec{v})=-\vec{v}^TA\vec{v}-\vec{\beta}^T\vec{v}
\end{equation*}

@@

\subsection{ESS strategies }
Surprisingly, even without imposing costs on paying attention to more neighbors, having as many neighbors as possible is usually sub-optimal. The ESS strategies for the number of neighbors to pay attention strongly depends on which selection pressure is being applied. The ESS strategy when selection is due to knowledge of a predator is always higher than the ESS strategy when selection is due to knowledge about resources (Figure \ref{ESS}). When selection is due to predation, the ESS strategy is a non-monotonic function of the the radius of the signal, i.e. how public the signal is (Figure \ref{ESS}). This non-monotonic behavior is due to two opposing effects of paying attention to more neighbors. First, having more neighbors increases the probability that one of them will have the signal and thus the probability that the focal bird will have a good source of information. Second, having more neighbors increases the probability that a focal bird will be trying to learn from individuals who do not have true information about the environment. When few individuals have the true signal (i.e. low radius $r$), the second effect outweighs the first and the ESS strategies are to pay attention to fewer birds than are in the whole flock. However, as more individuals have the true signal (i.e. higher radius $r$), the first effect is more important

@@

\section{Discussion}

\section{Figures }
\begin{figure}[ht]
\includegraphics[width=6.83in]{/Users/eleanorbrush/Desktop/ESSneighbors.pdf}
\caption{\label{ESS} The ESS number of neighbors is always higher if selection is due to predation than due to resources.  The ESS number of neighbors when selection is due to predation is a non-monotonic function of how public the signal is. The x-axis indicates the radius of the signal and the y-axis indicates the ESS number of neighbors. Parameters:  $\beta=1$, $T=1$. 
}
\end{figure}
